Data Analysis

### 2023-04-21 ms1_data_analysis.Rmd
Adding RDA position vs trait analyses.

## Experimental design

@@

@@

      print(paste(desc,st,trt))

@@

    print(paste(desc, trt, paste0(lbls, collapse = "")))

@@

Elevations were similar in their baseline productivity (total biomass). Mid and high elevations had the highest baseline productivity (total biomass). Abundance and species richness.  decreased with elevation. Control plots in Numba are more variable than in other two elevations. Yawan has one site with lowest abundance (garden number seven).However no significant differences in variance were detected (Levene's and Bartellet's tests)

## General effects of biotic factors

We calculated log-response ratios for plant biomass, species richness, species diversity and stem density to be able to compare the effect sizes at individual elevations.

Table 1. Test results for over-dispersion all descriptors at three studied elevations.

```{r, echo=FALSE, results='show', fig.show='all', warning=FALSE, message=FALSE}
sjPlot::tab_df(lrr_test_df)
```

@@

Table S2. Two models testing difference of LRR of predator exclosures from zero at each elevation.
```{r}
sjPlot::tab_df(model_comp)
```

There is generally little evidence for the over-dispersion in the LRR values at individual elevations. Only woody stem density for the predator exclusion showed some evidence of over-dispersion. Including this in statistical modle (using 'varIdent(form = ~1|site)' as weights in the GLS function) showed significant difference of LRR in Yawan, that was not observed for LM model.

@@

# Create a factor variable for sites
panel_data$fsite <- factor(panel_data$site, labels = c("Wanang",
                                      "Numba",
                                      "Yawan"))

@@

  theme_bw()+
  theme(strip.text = element_text(size =5.5),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1))+
  geom_hline(yintercept = 0, lty = 2)+
  ylab("Log response ratio")+xlab("")+
  facet_grid(descriptor ~ ftreatment)

@@

Positive effect of pathogenic fungi on richness was found at low and high elevations (Fig. 1) and on density at low elevation. Insects reduced biomass only at the mid-elevation, increased diversity at high and richness at high and low elevation. At high elevation predator removal caused an increase in diversity, richness and density of woody plants.Increase in generalist herbivore abundance reduced biomass and woody plant density at low elevation and richness at low and high elevations.

@@

## 3.1 Analyses for each elevation

plot.list <- list()

@@

@@

  p2 <- ggplot(species.dat, aes(x = RDA1*1.1, y = RDA2*1.1)) + 
    geom_text(data = species.dat %>%
                filter(names %in% sign.names), aes(label = names))+
    geom_segment(data = arrow.dat, aes(x = 0, y = 0, 
                                       xend = RDA1, yend = RDA2),
                 arrow = arrow(length=unit(0.30,"cm"), 
                               ends="last", 
                               type = "closed"),
                 alpha = 0.8, 
                 col = arr.col,
                 lwd = arr.lwd,
                 lty = arr.lty)+
    geom_text(data = arrow.dat, aes(x = RDA1*1.1, y = RDA2*1.1, 
                                    label = treats))+
    theme_minimal()

@@

# NEEDS TO BE FIXED

csites <- grepl("_c\\b", COMMVARS$plotid)
fsites <- grepl("_f", COMMVARS$plotid)
isites <- grepl("_i", COMMVARS$plotid)
hsites <- grepl("_h", COMMVARS$plotid)
psites <- grepl("_p", COMMVARS$plotid)

wsites <- COMMVARS$site == "wanang"
nsites <- COMMVARS$site == "numba"
ysites <- COMMVARS$site == "yawan"

@@

cdatc <- COMMDATA[grepl("c$", rownames(COMMDATA)), ]
cvars <- COMMVARS[rownames(cdatc), ]

# Make a matrix
# for(i in 1:dim(cdatc)[1]){
#   # Compare all other sites to the current i
#   for (j in (1:dim(cdatc)[1])[-i]) {
#     val <- beta.pair.abund(cdatc[c(i,j), ])
#   }
# }

# Comparisons between
wc.beta <- beta.pair.abund(COMMDATA[COMMVARS[wsites & csites,]$plotid,  ])

@@

for(siteid in c("w","y","n")){
  for(treat in c("_c\\b","_f","_h","_i","_p")){

@@

    print(rownames(site.treat.comm))

@@

    # Merge the datset
    beta.df <- rbind(beta.df,sub.df)

@@

# Fix the site order
beta.df$fsite <- factor(beta.df$site, levels = c("w","n","y"))

@@

# CLD data for comparisons between elevations
for (comp in unique(beta.df$component)){
  print(comp)
  for (trt in unique(beta.df$treat)){
    print(trt)

@@

    print(y_loc)
    print(sub.cld)

@@

# cld.dat = data.frame()
# prws.comp <- list()
# 
# # For data for comparisons between treatments we need pairwise comps 
# # (control vs treat) - this can be obtained from summary of the betareg
# 
# # library(glmmTMB)

@@

  for (ste in unique(beta.df$fsite)){
    print(comp)
    print(ste)

@@

pairwise.comp.res

# # Plot
# # tiff(filename='figures/beta_part_panel.tif',
# #      height=5600,
# #      width=5600,
# #      units='px',
# #      res=800,compression='lzw')

@@

# # tiff(filename='figures/beta_part_panel_treats.tif',
# #      height=5600,
# #      width=5600,
# #      units='px',
# #      res=800,compression='lzw')

# Alternative boxplot
beta.part.plot2 <- ggplot(beta.df, aes(x = treat, y = vals))+
  geom_boxplot()+
  # stat_summary(fun.y = mean, na.rm = T,
  #              geom = "point", size = 3, col = "gray30")+
  # stat_summary(fun.data = "mean_cl_boot", col = "gray30")+
  geom_text(data = pairwise.comp.res,
            aes(x = y_loc_trt, y = y_loc + 0.5,
                label = label), cex = 6,  col = "red", lwd = 2)+
  facet_grid(component ~ fsite) + 
  theme_bw()
```

```{r, fig.width=10, fig.height=18}
ggpubr::ggarrange(beta.part.plot1,
                  beta.part.plot2, nrow = 2)
```

@@

* Insect herbivores (P, I, H combined) can suppress plants in communities with low variability in species composition, which allows for new species to establish. Moreover, **insects reduce dissimilarity for highly variable communities** (is this true?). Exclusion of top predators has the same effect as insect exclusion on species turnover in Numba and Yawan. This is a probable result for a strong top-down control of the tri trophic interactions. Addition of generalist herbivores also correlated with the predator effects. If the effects of insects are mainly caused by specialists, then generalists interfering with their feeding could have the same results. But the effects can be more subtle than those of herbivore removal.

@@

# rfData <- lgsData %>%
#   mutate(pg.cat = as.factor(pg.cat),
#          lest.leaf.area = log(est.leaf.area),
#          lstem = log(no_stems)) %>%
#   dplyr::select(pg.cat, lstem, perc.herb, water.cont,
#                 lest.leaf.area, richness, ltb, lsla.d, treatment)
# 
# # Try standardized, non-correlated variables
# rfDataRF <- rfData %>%
#   filter(!is.infinite(lest.leaf.area)) %>%
#   dplyr::select(pg.cat, perc.herb, water.cont,
#                 richness, ltb, lsla.d)
# 
# # Standardized
# rfDataRF[,-1] <- scale(rfDataRF[,-1])
# 
# # Splitting data in train and test data
# split <- sample.split(rfDataRF, SplitRatio = 0.7)
# train <- subset(rfDataRF, split == "TRUE")
# test <- subset(rfDataRF, split == "FALSE")
#   
# classifier_RF = randomForest(x = train[-1],
#                              y = train$pg.cat,
#                              ntree = 500)

# classifier_RF

# Predicting the Test set results
# y_pred = predict(classifier_RF, newdata = test[-1])

# Variable importance plot
# varImpPlot(classifier_RF)
```

@@

```{r}
# CWM vs richness
```

@@

# Plot of CWM vs species richness
require(scales)
p1 <- ggplot(siteSLAdf, aes(x = sp.rich, y = cwm.sla, color = treatment))+
  geom_point(aes(size = total.bio)) + 
  scale_x_continuous(trans = 'log2',
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x))) +
  scale_y_continuous(trans = 'log2',
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x))) +  theme_minimal()

@@

```
```{r, results='asis'}
# Test for the relationship and the differences
require(sjPlot)
cwm.lm <- lm(log(cwm.sla)~log(sp.rich)*treatment + site, data = siteSLAdf)
cwm.lm.nosite <- lm(log(cwm.sla)~log(sp.rich)*treatment, data = siteSLAdf)
cwm.lm.nosite.noint <- lm(log(cwm.sla)~log(sp.rich)+treatment, data = siteSLAdf)
cwm.lm.onlysp <- lm(log(cwm.sla)~log(sp.rich), data = siteSLAdf)

@@

### Can SLA predict change in composition?

SLA doesn’t predict RDA position along C->I, C->H, C->P, C -> F axes, at any elevation.

# Figure

@@

Community composition change under a treatment may be independent of SLA. [key indicator of functionally important aspects of photosynthetic capacity, and is positively correlated with ecosystem productivity (White et al., 2000, Madani et al., 2017, Anderson et al., 2020). Reflects the abilities of light capture and adaption to environments of plants (Gao, Wang, and Zhang 2022)]

>>>>

# Experimental design

@@

# Results

@@

      # print(paste(desc,st,trt))

@@

    # print(paste(desc, trt, paste0(lbls, collapse = "")))

@@

Elevations were similar in their baseline productivity (total biomass). At the high elevation the plant abundance, richness and diversity was significantly higher than at the low elevation. Mid elevation has many outliers. However, no significant differences in variance were detected (Table S1)

## General effects of biotic factors
We calculated log-response ratios for plant biomass, species richness, species diversity and stem density to be able to compare the effect sizes at individual elevations.

Table 1. Test results for over-dispersion all descriptors at three studied elevations.
```{r, echo=FALSE, results='show', fig.show='all', warning=FALSE, message=FALSE}
sjPlot::tab_df(lrr_test_df)
```

@@

Table S2. Two models testing difference of LRR of predator exclosures from zero at each elevation. General Least Squares assumes over-dispersion in the data
```{r, results='asis'}
sjPlot::tab_df(model_comp)
```

There is generally no evidence for the over-dispersion in the LRR values at individual elevations. Only woody stem density for the predator exclusion showed some evidence of over-dispersion. Including this in statistical modle (using 'varIdent(form = ~1|site)' as weights in the GLS function) showed significant difference of LRR in Yawan, that was not observed for LM model.

@@

# Create a factor variable for sites
panel_data$fsite <- factor(panel_data$site, labels = c("Low",
                                      "Mid",
                                      "High"))

@@

  theme_bw()+
  theme(strip.text = element_text(size =5.5),
        axis.text.x = element_text(angle = 0,
                                   hjust = 0.5))+
  geom_hline(yintercept = 0, lty = 2)+
  ylab("Log response ratio")+xlab("")+
  facet_grid(descriptor ~ ftreatment)

@@

Positive effect of pathogenic fungi on richness was found at low and high elevations (Fig. 1) and on density at low elevation. Insects reduced biomass only at the mid-elevation, increased diversity at high and richness at high and low elevation. At high elevation predator removal caused an increase in diversity, richness and density of woody plants. Increase in generalist herbivore abundance reduced biomass and woody plant density at low elevation and richness at low and high elevations.

@@

## 3.1 Analyses for each elevation
ord.list <- list()
plot.list <- list()

@@

  ord.list[[site.id]] <- rdamod

@@

  p2 <- ggplot(species.dat, aes(x = RDA1*1.1, y = RDA2*1.1)) + 
    geom_text(data = species.dat %>%
                filter(names %in% sign.names), 
              aes(label = names, alpha = 0.2))+
    geom_segment(data = arrow.dat, aes(x = 0, y = 0, 
                                       xend = RDA1, yend = RDA2),
                 arrow = arrow(length=unit(0.30,"cm"), 
                               ends="last", 
                               type = "closed"),
                 alpha = 0.8, 
                 col = arr.col,
                 lwd = arr.lwd,
                 lty = arr.lty)+
    geom_text(data = arrow.dat, aes(x = RDA1*1.1, y = RDA2*1.1, 
                                    label = treats))+
    theme_minimal()

@@

```{r}
# csites <- grepl("_c\\b", COMMVARS$plotid)
# fsites <- grepl("_f", COMMVARS$plotid)
# isites <- grepl("_i", COMMVARS$plotid)
# hsites <- grepl("_h", COMMVARS$plotid)
# psites <- grepl("_p", COMMVARS$plotid)

# wsites <- COMMVARS$site == "wanang"
# nsites <- COMMVARS$site == "numba"
# ysites <- COMMVARS$site == "yawan"

@@

# cdatc <- COMMDATA[grepl("c$", rownames(COMMDATA)), ]
# cvars <- COMMVARS[rownames(cdatc), ]

# Comparisons between
# wc.beta <- beta.pair.abund(COMMDATA[COMMVARS[wsites & csites,]$plotid,  ])

@@

# siteid <- "^w"
# treat <- "_c\\b"

for(siteid in c("^w","^y","^n")){
  for(treat in c("_c\\b","_f","_h","_i","_p")){

@@

@@

    # Merge the data-set
    beta.df <- rbind(beta.df,sub.df)

@@

# Fix the site order
beta.df$fsite <- factor(beta.df$site, levels = c("^w","^n","^y"),
                        labels = c("w","n","y"))

@@

# CLD data for comparisons between elevations
for (comp in unique(beta.df$component)){
  for (trt in unique(beta.df$treat)){

@@

@@

@@

  for (ste in unique(beta.df$fsite)){
    # print(comp)
    # print(ste)

@@

@@

# Alternative boxplot
beta.part.plot2 <- ggplot(beta.df, aes(x = treat, y = vals))+
  geom_boxplot()+
  # stat_summary(fun.y = mean, na.rm = T,
  #              geom = "point", size = 3, col = "gray30")+
  # stat_summary(fun.data = "mean_cl_boot", col = "gray30")+
  geom_text(data = pairwise.comp.res,
            aes(x = y_loc_trt, y = y_loc + 0.5,
                label = label), cex = 6,  col = "red")+
  facet_grid(component ~ fsite) + 
  theme_bw()
```

```{r, fig.width=10, fig.height=16}
ggpubr::ggarrange(beta.part.plot1,
                  beta.part.plot2, nrow = 2)
```

@@

* Insect herbivores (P, I, H combined) can suppress plants in communities with low variability in species composition, which allows for new species to establish. Moreover, **insects reduce dissimilarity for highly variable communities** (is this true?). Exclusion of top predators has the same effect as insect exclusion on species turnover in Numba and Yawan. This is a probable result for a strong top-down control of the tri-trophic interactions. Addition of generalist herbivores also correlated with the predator effects. If the effects of insects are mainly caused by specialists, then generalists interfering with their feeding could have the same results. But the effects can be more subtle than those of herbivore removal.

@@

@@

@@

# Plot of CWM vs species richness
require(scales)
p1 <- ggplot(siteSLAdf, aes(x = sp.rich, y = cwm.sla))+
  geom_point(aes(size = total.bio, color = treatment)) + 
  stat_smooth(method = "lm")+
  scale_x_continuous(trans = 'log2',
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x))) +
  scale_y_continuous(trans = 'log2',
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x))) +  theme_minimal()

@@

```{r, results='asis'}
# Test for the relationship and the differences
require(sjPlot)
cwm.lm <- lm(log(cwm.sla)~log(sp.rich)*treatment + site, data = siteSLAdf)
cwm.lm.nosite <- lm(log(cwm.sla)~log(sp.rich)*treatment, data = siteSLAdf)
cwm.lm.nosite.noint <- lm(log(cwm.sla)~log(sp.rich)+treatment, data = siteSLAdf)
cwm.lm.onlysp <- lm(log(cwm.sla)~log(sp.rich), data = siteSLAdf)

@@

### Can traits predict change in composition?
SLA doesn’t predict RDA position along C->I, C->H, C->P, C -> F axes, at any elevation.
Community composition change under a treatment may be independent of SLA. [key indicator of functionally important aspects of photosynthetic capacity, and is positively correlated with ecosystem productivity (White et al., 2000, Madani et al., 2017, Anderson et al., 2020). Reflects the abilities of light capture and adaption to environments of plants (Gao, Wang, and Zhang 2022)]

```{r}
# Calculate position of species and site along a given treatment vector.
function (oridnation, treatment, type = c('site','species')) {
  NULL
}

surfData <- data.woody.brut %>%
  filter(site == "wanang") %>%
  group_by(unified_names) %>%
  summarise(lsla.d = log(mean(sla.d, na.rm=T)),
            water.cont = mean(water.cont, na.rm = T),
            herbivory = mean(perc.herb, na.rm = T))

# Add row names to sla data
surfDataDf <- as.data.frame(surfData)
tibble::column_to_rownames(surfData, var = 'unified_names')
rownames(surfDataDf) <- surfDataDf$unified_names

# Get the treatment vector.
nm <- "f"
# names(ord.list[["w"]])
trtVcts <- ord.list[["w"]]$CCA$biplot[nm, c(1,2)]

# Get species vector.
all.specs <- ord.list[["w"]]$CCA$v[, c(1,2)]
usednms <- surfData$unified_names[surfData$unified_names %in% rownames(all.specs)]
specVcts <- all.specs[usednms,]

# Projection of species onto treatment vector.
VdotU <- ((specVcts %*% trtVcts))
uNormSq <- (trtVcts %*% trtVcts)
projUV <- (VdotU/as.numeric(uNormSq)) %*% t(as.matrix(trtVcts))

# Get norm for each projection vector.
positions <- apply(projUV, 1, function(x) sqrt(x %*% x))

# Add positions to the data frame.
modDat <- data.frame(value = c(surfDataDf[usednms, ]$lsla.d,
                               surfDataDf[usednms, ]$water.cont,
                               surfDataDf[usednms, ]$herbivory),
                     label = rep(c("SLA", "Water Content", "Herbivory"), 
                                 each = length(positions)),
                     pos = positions[usednms])

testDat <- data.frame(sla = surfDataDf[usednms, ]$lsla.d,
                      water = surfDataDf[usednms, ]$water.cont,
                      herb = surfDataDf[usednms, ]$herbivory,
                      pos = positions[usednms])

ggplot(modDat, aes(x = pos, y = value)) +
  geom_point()+
  stat_smooth(method = "lm") + 
  facet_wrap(~label, scales = 'free')

# Tests + phylogeny!!!!
glsmod1 <- gls(pos ~ sla + water + herb, data = testDat)
glsmod2 <- gls(pos ~ sla +       + herb, data = testDat)
glsmod3 <- gls(pos ~ sla + water       , data = testDat)
glsmod4 <- gls(pos ~               herb, data = testDat)
glsmod5 <- gls(pos ~       water       , data = testDat)
glsmod6 <- gls(pos ~ sla               , data = testDat)

summary(glsmod1)
summary(glsmod2)
summary(glsmod3)
summary(glsmod4)
summary(glsmod5)
summary(glsmod6)

AIC(glsmod1,glsmod2,glsmod3,glsmod4,glsmod5,glsmod6)

summary(glsmod1)

myTree

# Correlate positions with individual species' with their sla. CWM we can leave out because we know it correlates with the .

@@
